---
title: "RQ_draft"
author: "Rui Qin"
date: "12/2/2020"
output: html_document
---

```{r}
df %>% 
  mutate(Time = factor(sprintf("%s:00",Hour))) %>%
  mutate(Time = fct_rev(Time)) %>%
  ggplot(aes(x=Time, y=num/365)) +
    geom_bar(stat="identity") +
    ylab("Number of riders") +
    coord_flip()
```
According to the graph above, it can be observed that the number of riders peaks during rush hours each day. Moreover, the number of riders gradually decreases and achieves its lowest point at 3:00 AM. 

```{r}
df %>% 
  group_by(Hour, Member.type) %>% 
  summarise(num=mean(num)) %>%
  mutate(Time = factor(sprintf("%s:00",Hour))) %>%
  mutate(Time = fct_rev(Time)) %>%
  ggplot(aes(x=Time, y=num)) + 
    geom_bar(stat="identity") +
    coord_flip() +
    facet_wrap(~Member.type)
```
When comparing number of riders between member and casual riders, we notice the number of casual riders is much lower than member riders. More importantly, there are different patterns of bike riding for two types of riders. For members, as in the last graph, the number of riders increases rapidly during rush hours. It can be predicted that many member riders use bike sharing for daily commute. On the other hand, for casual riders, most bike sharing happen during the daytime and there is no rapid increase in rush hours.

```{r}
df %>% 
  group_by(Month, Hour, Member.type) %>%
  summarise(num=mean(num)) %>%
  dplyr::filter(Month == "Jan") %>%
  mutate(Time = factor(sprintf("%s:00",Hour))) %>%
  mutate(Time = fct_rev(Time)) %>%
  mutate(num = ifelse(Member.type == "Member", num, -num)) %>%
  ggplot(aes(x=Time, y=num, fill=Member.type)) + 
    geom_bar(stat="identity") +
    coord_flip() +
    scale_y_continuous(name = "Number of riders", 
                       breaks = seq(-250,1000,250), 
                       limits = c(-250, 1000),
                       labels = abs(seq(-250,1000,250)))
```




```{r}
df %>% 
  group_by(Member.type, Hour) %>%
  summarise(num=mean(num)) %>%
  pivot_wider(names_from = Member.type, values_from = num) %>%
  select(Casual, Member)
```
```{r}
df %>% 
  group_by(Hour, Member.type) %>%
  summarise(num=mean(num))
```

```{r}
Q_1 = c("Jan", "Feb", "Mar")
Q_2 = c("Apr", "May", "Jun")
Q_3 = c("Jul", "Aug", "Sep")
Q_4 = c("Oct", "Nov", "Dec")
hour_inter_1 = df %>% 
  dplyr::filter(Month %in% Q_1) %>%
  group_by(Hour, Member.type) %>%
  summarise(num=mean(num)) %>%
  pivot_wider(names_from = Member.type, values_from = num) %>%
  select(Casual, Member)
hour_inter_2 = df %>% 
  dplyr::filter(Month %in% Q_2) %>%
  group_by(Hour, Member.type) %>%
  summarise(num=mean(num)) %>%
  pivot_wider(names_from = Member.type, values_from = num) %>%
  select(Casual, Member)
hour_inter_3 = df %>% 
  dplyr::filter(Month %in% Q_3) %>%
  group_by(Hour, Member.type) %>%
  summarise(num=mean(num)) %>%
  pivot_wider(names_from = Member.type, values_from = num) %>%
  select(Casual, Member)
hour_inter_4 = df %>% 
  dplyr::filter(Month %in% Q_4) %>%
  group_by(Hour, Member.type) %>%
  summarise(num=mean(num)) %>%
  pivot_wider(names_from = Member.type, values_from = num) %>%
  select(Casual, Member)
```


